### 2020-06-30 chapters/problem-statement.tex
Assumptions,  small corrections

\section{Proposal}
We devise a two-phased (batch and streaming phases) and two-windowed (reference and target sliding windows) method. We devise a sliding window approximate histogram aggregation that encodes data distributions and provide a method for incrementally maintaining it with constant time and space complexity for every growing volumes of data, ideal for the streaming scenario. Given a user-defined reference period and a stream of data to process, our method guarantees to monitor each event's fields and alert if they are considered divergent relative to their reference distribution. We provide a measure of how divergent each field (or feature) is. Based on the number of divergent features and the corresponding divergence measure, a system administrator has an easier time determining when to reconfigure the system.

@@

\begin{enumerate}[leftmargin=1.5cm, label=\textbf{(RQ\arabic*)}]
    \item What aggregation can we use to properly encode data distributions?
    \item How can we maintain such an aggregation in a sliding window fashion in real-time with constant memory usage?
    \item How can we measure divergence between the reference and sliding window aggregations?
    \item Based on a divergence measure, when should we raise an alert?
\end{enumerate}

%Assumptions section
%
%
%
%
%

>>>>

\section{Proposal}
We devise a two-phased (batch and streaming phases) and two-windowed (reference and target sliding windows) method. We devise a sliding window approximate histogram aggregation that encodes data distributions and provide a method for incrementally maintaining it with constant time and space complexity for every growing volumes of data, ideal for the streaming scenario. Given a user-defined reference period and a stream of data to process, we monitor each event's fields and alert if they are considered divergent relative to their reference distribution. We provide a measure of how divergent each field (or feature) is. Based on this information, it is our belief that a system administrator would have an easier time determining when to reconfigure the system.

\section{Assumptions}
Our feature monitoring method assumes only that the monitored features are numerical or categorical and that the events arrive in an orderly fashion during streaming. If the latter is not true, our method will still work but results may be imprecise, as our sliding window histogram aggregation relies on Exponential Moving Averages (EMAs), which give different weights to each event according to its age.

@@

\begin{enumerate}[leftmargin=1.75cm, label=\textbf{(RQ\arabic*)}]
    \item What aggregation can we use to properly encode data distributions?
    \item How can we maintain such an aggregation in a sliding window fashion in real-time with constant memory usage?
    \item How can we measure divergence between the reference and sliding window aggregations?
    \item Based on a divergence measure, when should we raise an alert?
\end{enumerate}

### 2020-06-30 chapters/validation.tex
Assumptions,  small corrections

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that combined covered a period of roughly 1 year and 9 months. We used the older one as the reference dataset and the more recent one as the target period. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus categorical features. Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & Tuesday, September 5, 2017 & Friday, June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & Friday, June 29, 2018      & Sunday, June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

@@

After the batch phase, we had the necessary inputs for the streaming phase (detailed in Section \ref{sec:batch-artifacts-summary}), namely, for each feature, the reference histograms, the list of observed divergence values and the last sample's histogram. We then used the target dataset, simulating a stream of ordered events. Our monitoring system processed the dataset event by event in sequential, time-ordered fashion, emulating a stream computation and reported divergent features and associated p-values. We performed our multiple hypotheses with the Holm-Bonferroni correction (Section \ref{sec:holmbonferroni}) for every 1000 events. It took us 382,363 milliseconds (roughly 6 minutes) to process the 4,032,505 transactions, amounting to a throughput of 10,546 transactions per second (TPS).

This merchant and in particular these datasets were known to suffer a lot from concept drift. Our system produced a large list of alerts, from nearly the beginning of the target period until the end. Due to the large number of features (183) we only show some plots. Furthermore, we refer to features with anonymized identifiers, like feature $x_t$. In the following plots, we show the corrected p-values (Formula \ref{for:corrected-pvalue}) for some features at each event of the time-series and the FWER threshold set at 1\%. A feature is considered divergent at a given timestamp if below that threshold line.

@@

Now let us look at features x2 and x3. Figures \ref{fig:merchant-x2-reference} and \ref{fig:merchant-x2-target} show feature x2' time-series plots for the reference and target periods, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-reference.png}
      \caption{Merchant feature x2 reference time-series}
      \label{fig:merchant-x2-reference}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-target.png}
      \caption{Merchant feature x2 target time-series}
      \label{fig:merchant-x2-target}
    \end{center}
\end{figure}
Figures \ref{fig:merchant-x3-reference} and \ref{fig:merchant-x3-target} show feature x3' time-series plots for the reference and target periods, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-reference.png}
      \caption{Merchant feature x3 reference time-series}
      \label{fig:merchant-x3-reference}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-target.png}
      \caption{Merchant feature x3 target time-series}
      \label{fig:merchant-x3-target}
    \end{center}
\end{figure}
\clearpage
Similarly to x1, features x2 and x3 have very similar reference and target periods (not between each other). Just like feature x1, we expect to have some alarms but still not too many.

Once again, to visualize the alarms, we plot the corrected p-values for each event of the target time-series for each feature. Figure \ref{fig:merchant-x2-correctedpvalues} is a plot of feature x2' corrected p-values and the FWER threshold of 1\%.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-correctedpvalues.png}
      \caption{Merchant feature x2 corrected p-values}
      \label{fig:merchant-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-correctedpvalues.png}
      \caption{Merchant feature x3 corrected p-values}
      \label{fig:merchant-x3-correctedpvalues}
    \end{center}
\end{figure}
The corrected p-values plot shows us once again that both x2 and x3 features had some alarms but quickly rose above the threshold and resumed their reference distribution. This is expected from features whose reference and target time-series plots look so similar.

@@

The -1 value is often used as placeholder for missing values, \textit{i.e.}, if some problem occurs and the value for a feature is missing at a given timestamp or event, -1 is used. A follow-up question that is raised is if we should include some mechanism of dealing with missing values in our method. However, we discard this idea and defend that we should not: missing values are indeed something we want to detect. For instance, if a certain device responsible for reporting one feature stops working, we want to know that. This scenario is very likely to happen in a distributed Internet-of-Things (IoT) system with hundreds of devices transmitting data.

%feats = [max_amt_rate_per_location_72h, num_preauths_per_card_1w, user_avg_amount_completions_1h, completions_rate_per_location_zip3_72h, completion_rate_per_card_72h]
%low drift:
    %x1 - max_amt_rate_per_location_72h
    %x2 - num_preauths_per_card_1w
    %x3 - user_avg_amount_completions_1h
%high drift:
    %x4 - completion_rate_per_card_72h
\subsection{Same reference and target periods}
In this next experiment, we set out to test our system in a low drift scenario,...

>>>>

Feedzai provides a fraud detection solution for this merchant and had stored two contiguous in time datasets of credit card transactions that combined covered a period of roughly 1 year and 9 months. In this experiment, we used the older one as the reference dataset and the more recent one as the target period. Both datasets were already feature-engineered \cite{Domingos-ML-Feat-Eng} and had a total of 183 numerical plus categorical features. Table \ref{tbl:merchant1-datasets-summary} shows summary statistics for both datasets.
\begin{table}[!htb]
    \begin{center}
        \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{Dataset} & \textbf{From}              & \textbf{To}           & \multicolumn{1}{l|}{\textbf{Transactions}} \\ \hline
        Reference        & Tuesday, September 5, 2017 & Friday, June 29, 2018 & 1,604,509                                  \\ \hline
        Target           & Friday, June 29, 2018      & Sunday, June 30, 2019 & 4,032,505                                  \\ \hline
        \end{tabular}
        \caption{Summary statistics for reference and target merchant datasets}
        \label{tbl:merchant1-datasets-summary}
    \end{center}
\end{table}

@@

After the batch phase, we had the necessary inputs for the streaming phase (detailed in Section \ref{sec:batch-artifacts-summary}), namely, for each feature, the reference histograms, the list of observed divergence values and the last sample's histogram. We then used the target dataset, simulating a stream of ordered events. Our monitoring system processed the dataset event by event in sequential, time-ordered fashion, emulating a stream computation and reported divergent features with associated p-values. We performed our multiple hypotheses with the Holm-Bonferroni correction (Section \ref{sec:holmbonferroni}) for every 1000 events. It took us 382,363 milliseconds (roughly 6 minutes) to process the 4,032,505 transactions, amounting to a throughput of 10,546 transactions per second (TPS).

This merchant and in particular these datasets were known to suffer a lot from concept drift. Our system produced a large list of alerts, from nearly the beginning of the target period until the end. Due to a large number of features (183), we only show some plots. Furthermore, we refer to features with anonymized identifiers, like feature $x_t$. In the following plots, we show the corrected p-values (Formula \ref{for:corrected-pvalue}) for some features at each event of the time-series and the FWER threshold set at 1\%. A feature is considered divergent at a given timestamp if below that threshold line.

@@

Now let us look at features x2 and x3. Figures \ref{fig:merchant-x2-reference} and \ref{fig:merchant-x2-target} show feature x2' time-series plots for the reference and target periods, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-reference.png}
      \caption{Merchant feature x2 reference time-series}
      \label{fig:merchant-x2-reference}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-target.png}
      \caption{Merchant feature x2 target time-series}
      \label{fig:merchant-x2-target}
    \end{center}
\end{figure}
Figures \ref{fig:merchant-x3-reference} and \ref{fig:merchant-x3-target} show feature x3' time-series plots for the reference and target periods, respectively.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-reference.png}
      \caption{Merchant feature x3 reference time-series}
      \label{fig:merchant-x3-reference}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-target.png}
      \caption{Merchant feature x3 target time-series}
      \label{fig:merchant-x3-target}
    \end{center}
\end{figure}
\clearpage
Similarly to x1, features x2 and x3 have very similar reference and target periods (not between each other). Just like feature x1, we expect to have some alarms but still not too many. Once again, to visualize the alarms, we plot the corrected p-values for each event of the target time-series for each feature. Figure \ref{fig:merchant-x2-correctedpvalues} is a plot of feature x2' corrected p-values and the FWER threshold of 1\%. Likewise, Figure \ref{fig:merchant-x3-correctedpvalues} shows the same plot but for feature x3.
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x2-correctedpvalues.png}
      \caption{Merchant feature x2 corrected p-values}
      \label{fig:merchant-x2-correctedpvalues}
    \end{center}
\end{figure}
\begin{figure}[!htb]
    \begin{center}
      \includegraphics[scale=0.5]{figures/merchant-x3-correctedpvalues.png}
      \caption{Merchant feature x3 corrected p-values}
      \label{fig:merchant-x3-correctedpvalues}
    \end{center}
\end{figure}
The corrected p-values plot shows us once again that both x2 and x3 features had some alarms but quickly rose above the threshold and resumed their reference distribution. This is expected from features whose reference and target time-series plots look so similar.

@@

The -1 value is often used as a placeholder for missing values, \textit{i.e.}, if some problem occurs and the value for a feature is missing at a given timestamp or event, -1 is used. A follow-up question that is raised is if we should include some mechanism of dealing with missing values in our method. However, we discard this idea and defend that we should not: missing values are indeed something we want to detect. For instance, if a certain device responsible for reporting one feature stops working, we want to know that. This scenario is likely to happen in a distributed Internet-of-Things (IoT) system with hundreds of devices transmitting data.

%feats = [max_amt_rate_per_location_72h, num_preauths_per_card_1w, user_avg_amount_completions_1h, completions_rate_per_location_zip3_72h, completion_rate_per_card_72h]
%low drift:
    %x1 - max_amt_rate_per_location_72h
    %x2 - num_preauths_per_card_1w
    %x3 - user_avg_amount_completions_1h
%high drift:
    %x4 - completion_rate_per_card_72h
\subsection{Same reference and target periods}
In this experiment, we set out to test our system in a low drift scenario. To guarantee the reference and target periods are as similar as possible, we decide to use the same dataset to work as reference and target. We chose to use the reference dataset from the previous experiment (Reference of Table \ref{tbl:merchant1-datasets-summary}) and feed it as input for both our batch and stream analysis as the reference and target periods, respectively.

\textcolor{red}{TODO: visualizations and conclusions of this experiment}
